---
title: "Stat558_Project3_CartronKaranjai"
author: "Matthieu Cartron and Sneha Karanjai"
date: "11/11/2022"
output: html_document
---

# Setup

Note: to hide later if necessary.

## Libraries

```{r lib, message=FALSE, warning=FALSE, error=FALSE}
library(tidyverse)
library(caret)
```

# Introduction

# Data

```{r readingdata, message=FALSE, warning=FALSE}
unzippedNewDataCSV <- unzip("OnlineNewsPopularity.zip")

newsDataName <- read_csv(unzippedNewDataCSV[1]) # This is the names file
newsData <- read_csv(unzippedNewDataCSV[2])

head(newsData)
```

```{r}
newsData %>% 
  select(starts_with("data_channel_is_"))
```
The 6 groups to analyze are : 

- Lifestyle 
- Entertainment 
- Business
- Social Media 
- Technology 
- World 

We will filter the data to analyze articles in one data channel at a time. Additionally according to the data report, `url` and `timedelta` are two non-predictive columns so we will remove them.

```{r subsetdata}
subsettingData <- function(data, area){
  #getting the naming convention as per the dataframe
  key <- ifelse(tolower(area)=="lifestyle", "lifestyle",
                ifelse(tolower(area)=="entertainment", "entertainment",
                       ifelse(tolower(area)=="business", "bus",
                              ifelse(tolower(area)=="social media", "socmed",
                                     ifelse(tolower(area)=="technology", "tech",
                                            ifelse(tolower(area)=="world", "world", "NA"))))))
  subsetVar <- paste("data_channel_is_", key, sep = "")
  
  # filtering the data and removing the data_channel_is_ columns, url, and timedelta
  subsetData <- data %>% 
    filter(!!as.symbol(subsetVar)==1) %>% 
    select(-c(starts_with("data_channel_is_"), url, timedelta))
  
  return(subsetData)
}

data <- subsettingData(newsData, "lifestyle")
```

# Exploratory Data Analysis 

Let us take a look at the columns available. 

```{r}
colnames(data)
```

Whew! That is a long list of columns to analyze. #Running a quick summary statistics on them.# Instead of analyzing them all, let us think about our data. What might we expect to be related to how many times an article is shared? We hear frequently about how the dissemination of news and the content thereof are related in some way. For our data channels, let's pay close attention to the number of shares (dissemination) and variables that we might be able to link to it. Can we find any interesting relationships in the exploratory analysis?  

```{r}
head(data)
```


```{r}

ggpairs(data, columns = c(43, 46, 50, 53))

```


